The locally Gaussian partial correlation

### 2020-09-16 responses/response_ae.Rmd

\noindent {Here are our responses to the comments from the Associate Editor:}

\vspace{1cm}

**AE’s general comment:**  I have heard from two referees. The views are mixed. Referee 2  is not enthusiastic about the paper and recommends rejection while referee 1 recommends a major revision. I also read the paper carefully and understands both referees concerns

> **Our response:** Thank you for reading our paper so carefully. We are very grateful for your constructive comments which have lead to a much improved paper.

@@

In addition we have many available tests for conditional independence between a vector and scalar variables (and also between two vectors of random variables) conditional on a third vector random variable. However, the test proposed in the paper is limited to testing conditional independence between two scalars only. In addition, the simulation results are definitely not better than the existing tests.

Also, I believe the asymptotic and bootstrap theories are not complete. I also made other comments.

For the above reasons, I believe that the contributions in this paper is definitely below the level of JBES.

> **Our response:** We have made a response directly to the 6 points in the report of referee 2. We briefly summarize this below, at the same time responding to the cover letter of referee 2: \begin{itemize}
\item[1.] \textbf{The novelty of the paper.} We look as the main contribution of our paper the introduction of the new concept of local partial Gaussian correlation (LGPC). With this concept we have sought to make a nonlinear and local extension of the ordinary (global) partial correlation, whose use in measuring and describing dependence is well known for Gaussian distributions. We claim that it is fundamentally different, both in concept and in use, from the existing tests of conditional independence. It is hard to obtain a measure of conditional dependence that can be easily interpreted from these tests as it would have to be based on distance measures that for example cannot distinguish between negative and positive conditional dependence.
\vspace{.3cm} \newline 
Using the kernel estimates for the purpose of measuring conditional dependence is not easy in the same way as it is not easy to use the joint distribution function for a numeric measure of dependence. See item 1 in our response to referee 2. 
\vspace{.3cm} \newline 
Clearly we did not get our message concerning our main contribution through to this referee. In retrospect we see that this may be our own fault. We did not succeed in adequately presenting the main purpose of our paper. Our main message is not that of proposing just another test of conditional independence,
and we have tried to make this more clear in the present version. Our test did do well in a competition with other test, and it may be possible to have it improved with better focusing and use of more local parameters, but, at the risk of overstating this, the main contribution, we believe, is the introduction and use of the new concept of LGPC in measuring the conditional dependence by considering a nonlinear local version of the partial correlation, a measure that reduces to the ordinary partial correlation for Gaussian variables.
\item [2.] \textbf{Estimation and testing in the full vector case:} 
We have developed a theory for this. More details can be found in the the response to item 3 in the report of referee 2. We have also added a new example where $X_1$ and $X_2$ are not both scalars.
\item [3.] \textbf{Asymptotic theory and bootstrap validity}. 
This has been developed in Sections E, F and H in the supplement, and details can be found in our response to item 4 in the report of referee 2. Actually, we are happy that referee 2 raised this point. We believe that it has strengthened the paper considerably. We add that we have also examined local power and the Pitman criterion in Section  G of the supplement. See also our response to item 5 of referee 2.
\end{itemize}

**AE's further general comment:** Referee 1 enjoys reading the paper and have some comments that can be relatively easily addressed.

@@

> **Our response:** We are grateful for the opportunity to revise, and we have responded to all of the comments of the two referees, and we will respond to your comments below.

**Comment 1:** I like the idea to extend the local likelihood framework of Hjort and Jones (1996) to estimate the local correlation and to extend the measure of local Gaussian correlation of Tjøstheim and Hufthammer (2013) to the measure of local Gaussian partial correlation (LGPC).

> **Our response:** Thank you.

**Comment 2:** As both referees mention (see R1's point 1 and R2's point 3), one limitation of the proposed testing procedure is that it only tests the conditional independence between two scalar variables $(X_1,X_2)$ given a random vector $(X_3,\ldots,X_p)$. It is important to generalize the test to allow
both $X_1$ and $X_2$ to be random vectors.

> **Our response:** Thank you for this important input. We now introduce and define the LGPC (matrix) for a general ${\bf X}_1$, ${\bf X}_2$ and ${\bf X}_3$ of dimension $d_1$, $d_2$ and $d_3$ in Section 2.1. The full multivariate framework is analyzed in more detail in Section A: "The multivariate LPGC" in the supplement. Various special cases are discussed in Section A.2 of that supplement. There are really no new concepts involved, and large parts of the ensuing estimation and testing theory in the supplement are carried out in this framework. 
\vspace{.3cm} \newline 
In particular the asymptotic theory for a conditional independence test where both ${\bf X}_1$ and ${\bf X_2}$ (and ${\bf X}_3$) are vectors is given in Section F of the supplementary material (Theorem F.1 and Theorem F.2). The asymptotic local power is discussed in Section G, and the validity of the bootstrap for the testing statistic are given at this level of generality (Theorems H.1 and H.2). However, the notation becomes more complex and expressions less explicit, and for these reasons and for reasons of space much of the material in the main article is stated for the case of a scalar $X_1$ and $X_2$. There is a new example of Granger causality in Section 4, though, where the assumption of a scalar $X_1$ and $X_2$ is dropped. The curse of dimensionality can be avoided at the cost of a simplified pairwise model introduced, described and used in the main article and in the supplement. Throughout the main paper the reader is referred to appropriate sections of the supplement for the general vector case.

@@

**Comment 3:** p.5: As Referee 1 points out, the density in (6) is not the real density of ${\Z}$. It is a local Gaussian approximation.

> **Our response:** You are certainly right. This was not well written. We have changed "representation" to "approximation" and also linked this directly to the defining approximative equation (4).

**Comment 4.** p.10: Since you claim that ${\bf b}$ is a diagonal matrix of bandwidths (not squared bandwidths). It seems that you want to define $K_{\bf b} = |{\bf b}|^{-1}K({\bf b}^{-1}{\bf x})$.

> **Our response:** Yes, corrections have been implemented. Thank you.

@@

\item [(d)] The results are too limited as it is only for the case with $p=3$. A general result for the generic $p$ is needed.

\item [(e)] Theorem 3.1 is a pointwise result. For the inference purpose, one typically needs a uniform result.

\end{itemize}

> **Our response:**
\begin{itemize}
\item[(a)] The reason for the slower convergence can basically be found in the expressions for ${\bf J}_{\bf b}$ and ${\bf M}_{\bf b}$. These expressions contain the matrix ${\bf u}(\cdot) {\bf u}^T(\cdot)$. This matrix has rank 1, and as $b \to 0$ this implies that the matrices ${\bf J}_{\bf b}$ and ${\bf M}_{\bf b}$ are singular and of order $O(b^2)$. It is therefore necessary to do a Taylor expansion to find the correct scaling factor, and this in turn leads to a the stated convergence rate of $(nb^5)^{-1/2}$. (Note that there was a typo in eq. (20). It has now been corrected.) This is explained in the proof of Theorem 3.1 and in more detail in Section 4 of Tjøstheim and Hufthammer (2013)  in  a slightly different situation. Similar results and with  rather different derivations can also be found in Jones (1996, 1998). In the pairwise situation referred to after the statement of Theorem 3.1, the matrices ${\bf J}_{\bf b}$ and ${\bf M}_{\bf b}$ collapse to scalars, the singularity disappears and one is back to the usual rates for the case in nonparametric kernel estimation with the dimension of the conditioning variable determining the convergence rate.
\vspace{.3cm} \newline It is certainly possible to estimate the local conditional correlation (and in the general vector case) using local likelihood for local correlations and then transforming to the LGPC. In the general vector case this is explained in Sections A.2 and A.3 in the supplement.
\item[(b)] This has been corrected. Thank you.
\item[(c)] This has been corrected. Thank you.
\item[(d)] Agree and the general vector case has now been implemented. We refer to our response to your comment 2 above.
\item[(e)] We agree that our results are pointwise results, thus resulting in pointwise confidence intervals for example. We have to leave uniform results and confidence bands to future investigations.
\end{itemize}

@@

**Comment 6:** p.16: Is there a data-driven choice of bandwidth? The choice of $c=4$ seems too arbitrary.

> **Our response:** There is a data-driven choice for bandwidth in a density estimation context in Otneim and Tjøstheim (2017). It is based on leave-one-out cross-validation. This is referred to in the beginning of Section 4. It remains to establish such a procedure for the present situation. In the paper we have typically used the asymptotic rates $b=cn^{-1/9}$ (with $c$ empirically determined to 1.75 in the density estimation context) for the full trivariate fit and $b=cn^{-1/6}$ for the bivariate simplification. In the testing results (Table 2, $2^*$ and $3^*$) we have used $c=1.0$ and $c=1.4$ to examine sensitivity to bandwidth selection. They gave similar results. Note also that for some of the other competing tests several choices of tuning parameters were used. For the plots of the estimated LGPC in Section 4 it is true that we used more smoothing to obtain cleaner plots. Thus $c=4$ for Figure 1, but a smaller bandwidth is used in testing and as mentioned in the description of that example the hypothesis of conditional independence is not rejected for any reasonable significance level.

@@

**Comment 7:** p.24-26: The authors propose a new test for conditional independence based on the LGPC. The test statistics should be formally studied under the null hypothesis of conditional independence and the usual Pitman sequence of local alternatives. For this purpose you need uniform consistent estimate of the local correlations. As referee 2 mentions in his/her fourth comment, you also need to justify the asymptotic validity of the bootstrap.

> **Our response:** The asymptotic properties of the test statistic has been formally studied in Section F of the supplement. The validity of the bootstrap for the test statistic is established in Section H of the supplement (for estimation in Section E of the supplement. The convergence rate of the test statistic is given in Section 5.2 in the main article. It is $n^{-1/2}$ for the pairwise testing case and $(nb^2)^{-1/2}$ for the trivariate testing case. This is briefly compared to rates for other test statistics. In Section G we look at local asymptotic  power and the Pitman criterion. We essentially use the approach of Wang and Hong (2018). (They avoid the uniformity argument presumably by using the weaker condition $D^{\prime}$ in Noether, G.E. (1955).)

@@

**Comment 8:** p.26: For the comparison with existing tests, you need to compare asymptotic power properties too.

> **Our response:** We refer to our response to comment 7, where we have compared convergence rates and indirectly local power

**Comment 9:** The paper contains a lot of typos. Referee 1 pointed out some of them and the authors should correct others too.

> **Our response:** We have corrected the typos kindly pointed out to us by Referee 1. We have read carefully through the revised version, including the supplement in order to eliminate as many typos as possible.

### 2020-09-16 responses/response_ae.tex

\noindent {Here are our responses to the comments from the Associate Editor:}

\vspace{1cm}

\textbf{AE's general comment:} I have heard from two referees. The views are mixed. Referee 2 is not enthusiastic about the paper and recommends rejection while referee 1 recommends a major revision. I also read the paper carefully and understands both referees concerns

\begin{quote}
\textbf{Our response:} Thank you for reading our paper so carefully. We are very grateful for your constructive comments which have lead to a much improved paper.
\end{quote}

@@

Also, I believe the asymptotic and bootstrap theories are not complete. I also made other comments.

For the above reasons, I believe that the contributions in this paper is definitely below the level of JBES.

\begin{quote}
\textbf{Our response:} We have made a response directly to the 6 points in the report of referee 2. We briefly summarize this below, at the same time responding to the cover letter of referee 2:

\begin{itemize}
\item[1.] \textbf{The novelty of the paper.} We look as the main contribution of our paper the introduction of the new concept of local partial Gaussian correlation (LGPC). With this concept we have sought to make a nonlinear and local extension of the ordinary (global) partial correlation, whose use in measuring and describing dependence is well known for Gaussian distributions. We claim that it is fundamentally different, both in concept and in use, from the existing tests of conditional independence. It is hard to obtain a measure of conditional dependence that can be easily interpreted from these tests as it would have to be based on distance measures that for example cannot distinguish between negative and positive conditional dependence.
\vspace{.3cm} \newline 
Using the kernel estimates for the purpose of measuring conditional dependence is not easy in the same way as it is not easy to use the joint distribution function for a numeric measure of dependence. See item 1 in our response to referee 2. 
\vspace{.3cm} \newline 
Clearly we did not get our message concerning our main contribution through to this referee. In retrospect we see that this may be our own fault. We did not succeed in adequately presenting the main purpose of our paper. Our main message is not that of proposing just another test of conditional independence,
and we have tried to make this more clear in the present version. Our test did do well in a competition with other test, and it may be possible to have it improved with better focusing and use of more local parameters, but, at the risk of overstating this, the main contribution, we believe, is the introduction and use of the new concept of LGPC in measuring the conditional dependence by considering a nonlinear local version of the partial correlation, a measure that reduces to the ordinary partial correlation for Gaussian variables.
\item [2.] \textbf{Estimation and testing in the full vector case:} 
We have developed a theory for this. More details can be found in the the response to item 3 in the report of referee 2. We have also added a new example where $X_1$ and $X_2$ are not both scalars.
\item [3.] \textbf{Asymptotic theory and bootstrap validity}. 
This has been developed in Sections E, F and H in the supplement, and details can be found in our response to item 4 in the report of referee 2. Actually, we are happy that referee 2 raised this point. We believe that it has strengthened the paper considerably. We add that we have also examined local power and the Pitman criterion in Section  G of the supplement. See also our response to item 5 of referee 2.
\end{itemize}
\end{quote}

\textbf{AE's further general comment:} Referee 1 enjoys reading the paper and have some comments that can be relatively easily addressed.

@@

\begin{quote}
\textbf{Our response:} We are grateful for the opportunity to revise, and we have responded to all of the comments of the two referees, and we will respond to your comments below.
\end{quote}

\textbf{Comment 1:} I like the idea to extend the local likelihood framework of Hjort and Jones (1996) to estimate the local correlation and to extend the measure of local Gaussian correlation of Tjøstheim and Hufthammer (2013) to the measure of local Gaussian partial correlation (LGPC).

\begin{quote}
\textbf{Our response:} Thank you.
\end{quote}

\textbf{Comment 2:} As both referees mention (see R1's point 1 and R2's point 3), one limitation of the proposed testing procedure is that it only tests the conditional independence between two scalar variables \((X_1,X_2)\) given a random vector \((X_3,\ldots,X_p)\). It is important to generalize the test to allow
both \(X_1\) and \(X_2\) to be random vectors.

\begin{quote}
\textbf{Our response:} Thank you for this important input. We now introduce and define the LGPC (matrix) for a general \({\bf X}_1\), \({\bf X}_2\) and \({\bf X}_3\) of dimension \(d_1\), \(d_2\) and \(d_3\) in Section 2.1. The full multivariate framework is analyzed in more detail in Section A: ``The multivariate LPGC'' in the supplement. Various special cases are discussed in Section A.2 of that supplement. There are really no new concepts involved, and large parts of the ensuing estimation and testing theory in the supplement are carried out in this framework.
\vspace{.3cm} \newline 
In particular the asymptotic theory for a conditional independence test where both \({\bf X}_1\) and \({\bf X_2}\) (and \({\bf X}_3\)) are vectors is given in Section F of the supplementary material (Theorem F.1 and Theorem F.2). The asymptotic local power is discussed in Section G, and the validity of the bootstrap for the testing statistic are given at this level of generality (Theorems H.1 and H.2). However, the notation becomes more complex and expressions less explicit, and for these reasons and for reasons of space much of the material in the main article is stated for the case of a scalar \(X_1\) and \(X_2\). There is a new example of Granger causality in Section 4, though, where the assumption of a scalar \(X_1\) and \(X_2\) is dropped. The curse of dimensionality can be avoided at the cost of a simplified pairwise model introduced, described and used in the main article and in the supplement. Throughout the main paper the reader is referred to appropriate sections of the supplement for the general vector case.
\end{quote}

@@

\textbf{Comment 3:} p.5: As Referee 1 points out, the density in (6) is not the real density of \({\Z}\). It is a local Gaussian approximation.

\begin{quote}
\textbf{Our response:} You are certainly right. This was not well written. We have changed ``representation'' to ``approximation'' and also linked this directly to the defining approximative equation (4).
\end{quote}

\textbf{Comment 4.} p.10: Since you claim that \({\bf b}\) is a diagonal matrix of bandwidths (not squared bandwidths). It seems that you want to define \(K_{\bf b} = |{\bf b}|^{-1}K({\bf b}^{-1}{\bf x})\).

\begin{quote}
\textbf{Our response:} Yes, corrections have been implemented. Thank you.
\end{quote}

@@

\item [(e)] Theorem 3.1 is a pointwise result. For the inference purpose, one typically needs a uniform result.

\end{itemize}

\begin{quote}
\textbf{Our response:}

\begin{itemize}
\item[(a)] The reason for the slower convergence can basically be found in the expressions for ${\bf J}_{\bf b}$ and ${\bf M}_{\bf b}$. These expressions contain the matrix ${\bf u}(\cdot) {\bf u}^T(\cdot)$. This matrix has rank 1, and as $b \to 0$ this implies that the matrices ${\bf J}_{\bf b}$ and ${\bf M}_{\bf b}$ are singular and of order $O(b^2)$. It is therefore necessary to do a Taylor expansion to find the correct scaling factor, and this in turn leads to a the stated convergence rate of $(nb^5)^{-1/2}$. (Note that there was a typo in eq. (20). It has now been corrected.) This is explained in the proof of Theorem 3.1 and in more detail in Section 4 of Tjøstheim and Hufthammer (2013)  in  a slightly different situation. Similar results and with  rather different derivations can also be found in Jones (1996, 1998). In the pairwise situation referred to after the statement of Theorem 3.1, the matrices ${\bf J}_{\bf b}$ and ${\bf M}_{\bf b}$ collapse to scalars, the singularity disappears and one is back to the usual rates for the case in nonparametric kernel estimation with the dimension of the conditioning variable determining the convergence rate.
\vspace{.3cm} \newline It is certainly possible to estimate the local conditional correlation (and in the general vector case) using local likelihood for local correlations and then transforming to the LGPC. In the general vector case this is explained in Sections A.2 and A.3 in the supplement.
\item[(b)] This has been corrected. Thank you.
\item[(c)] This has been corrected. Thank you.
\item[(d)] Agree and the general vector case has now been implemented. We refer to our response to your comment 2 above.
\item[(e)] We agree that our results are pointwise results, thus resulting in pointwise confidence intervals for example. We have to leave uniform results and confidence bands to future investigations.
\end{itemize}
\end{quote}

@@

\textbf{Comment 6:} p.16: Is there a data-driven choice of bandwidth? The choice of \(c=4\) seems too arbitrary.

\begin{quote}
\textbf{Our response:} There is a data-driven choice for bandwidth in a density estimation context in Otneim and Tjøstheim (2017). It is based on leave-one-out cross-validation. This is referred to in the beginning of Section 4. It remains to establish such a procedure for the present situation. In the paper we have typically used the asymptotic rates \(b=cn^{-1/9}\) (with \(c\) empirically determined to 1.75 in the density estimation context) for the full trivariate fit and \(b=cn^{-1/6}\) for the bivariate simplification. In the testing results (Table 2, \(2^*\) and \(3^*\)) we have used \(c=1.0\) and \(c=1.4\) to examine sensitivity to bandwidth selection. They gave similar results. Note also that for some of the other competing tests several choices of tuning parameters were used. For the plots of the estimated LGPC in Section 4 it is true that we used more smoothing to obtain cleaner plots. Thus \(c=4\) for Figure 1, but a smaller bandwidth is used in testing and as mentioned in the description of that example the hypothesis of conditional independence is not rejected for any reasonable significance level.
\end{quote}

@@

\textbf{Comment 7:} p.24-26: The authors propose a new test for conditional independence based on the LGPC. The test statistics should be formally studied under the null hypothesis of conditional independence and the usual Pitman sequence of local alternatives. For this purpose you need uniform consistent estimate of the local correlations. As referee 2 mentions in his/her fourth comment, you also need to justify the asymptotic validity of the bootstrap.

\begin{quote}
\textbf{Our response:} The asymptotic properties of the test statistic has been formally studied in Section F of the supplement. The validity of the bootstrap for the test statistic is established in Section H of the supplement (for estimation in Section E of the supplement. The convergence rate of the test statistic is given in Section 5.2 in the main article. It is \(n^{-1/2}\) for the pairwise testing case and \((nb^2)^{-1/2}\) for the trivariate testing case. This is briefly compared to rates for other test statistics. In Section G we look at local asymptotic power and the Pitman criterion. We essentially use the approach of Wang and Hong (2018). (They avoid the uniformity argument presumably by using the weaker condition \(D^{\prime}\) in Noether, G.E. (1955).)
\end{quote}

@@

\textbf{Comment 8:} p.26: For the comparison with existing tests, you need to compare asymptotic power properties too.

\begin{quote}
\textbf{Our response:} We refer to our response to comment 7, where we have compared convergence rates and indirectly local power
\end{quote}

\textbf{Comment 9:} The paper contains a lot of typos. Referee 1 pointed out some of them and the authors should correct others too.

\begin{quote}
\textbf{Our response:} We have corrected the typos kindly pointed out to us by Referee 1. We have read carefully through the revised version, including the supplement in order to eliminate as many typos as possible.
\end{quote}

### 2020-09-16 responses/response_editor.Rmd

---
title: " "
output: 
  bookdown::pdf_document2:
      dev: tikz
      keep_tex: yes
      number_sections: yes
      toc: no
      includes:
          in_header: latex_preamble_supp.tex
classoption: letterpaper
geometry: margin = 1in
biblio-style: "apalike"
link-citations: yes
indent: yes
fontsize: 12pt
---

\noindent \textbf{Letter to the Editor.}

\noindent Professor Jianqing Fan

\noindent Editor JBES

@@

\noindent Dear Professor Fan.

\vspace{1cm}

We refer to your letter of 12 February 2020 where we are invited to revise our manuscript JBES-P-2019-0538, "The locally Gaussian partial correlation".

We grateful for the opportunity to revise. Enclosed please find our revised version, including supplementary material to be posted online, as well as individual responses to the Associate Editor and two referees.

@@

We apologize for the somewhat long time taken to complete the revision. This is due to the implementations of a rather large number of changes prompted by the detailed and constructive comments of reviewers. We thank them for their positive attitude to our paper and for reading the paper so carefully.

We believe that we have taken all of the queries of the referees and the Associate Editor into account, and in the accompanying files we have made point by point responses. As a result of the input received we think that the paper is much improved. Here are the major changes and improvements:

The local Gaussian partial correlation (LPGC) has been introduced and analyzed in a general vector case looking at the local joint conditional distribution of two random vectors $\X_1$ and $\X_2$ of dimension $d_1$ and $d_2$, respectively, given a third vector $\X_3$ of dimension $d_3$. An example concerning Granger causality testing where $\X_1$ and $\X_2$ are not scalars has been added. We have added asymptotic theory for the test statistic, theory on the validity of the bootstrap for the test statistic, and for the estimation of LGPC. We have also derived local power and compared with other test statistics.

For reasons of space most of these changes have been included in a supplement intended to be available online. This has made it possible for us to keep the length of the main article at 35 pages, as advised by the Editor.

In addition there are numerous other changes directly related to the comments of the referees and the Associate Editor. In particular,  we have tried to make it more clear that our paper is not just concerned with presenting just another test of conditional independence. But rather it is concerned with the introduction of a new framework for measuring conditional dependence, where this framework reduces to the ordinary partial correlation coefficient in the Gaussian case.

### 2020-09-16 responses/response_editor.tex

\definecolor{Gray}{gray}{0.9}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\noindent \textbf{Letter to the Editor.}

\noindent Professor Jianqing Fan

\noindent Editor JBES

@@

We grateful for the opportunity to revise. Enclosed please find our revised version, including supplementary material to be posted online, as well as individual responses to the Associate Editor and two referees.

We apologize for the somewhat long time taken to complete the revision. This is due to the implementations of a rather large number of changes prompted by the detailed and constructive comments of reviewers. We thank them for their positive attitude to our paper and for reading the paper so carefully.

We believe that we have taken all of the queries of the referees and the Associate Editor into account, and in the accompanying files we have made point by point responses. As a result of the input received we think that the paper is much improved. Here are the major changes and improvements:

The local Gaussian partial correlation (LPGC) has been introduced and analyzed in a general vector case looking at the local joint conditional distribution of two random vectors \(\X_1\) and \(\X_2\) of dimension \(d_1\) and \(d_2\), respectively, given a third vector \(\X_3\) of dimension \(d_3\). An example concerning Granger causality testing where \(\X_1\) and \(\X_2\) are not scalars has been added. We have added asymptotic theory for the test statistic, theory on the validity of the bootstrap for the test statistic, and for the estimation of LGPC. We have also derived local power and compared with other test statistics.

For reasons of space most of these changes have been included in a supplement intended to be available online. This has made it possible for us to keep the length of the main article at 35 pages, as advised by the Editor.

In addition there are numerous other changes directly related to the comments of the referees and the Associate Editor. In particular, we have tried to make it more clear that our paper is not just concerned with presenting just another test of conditional independence. But rather it is concerned with the introduction of a new framework for measuring conditional dependence, where this framework reduces to the ordinary partial correlation coefficient in the Gaussian case.

### 2020-09-18 lgpc_JBES_blinded.Rmd
Created blinded versions and zipped extra material

---
title: "The locally Gaussian partial correlation"
author: Håkon Otneim^[Corresponding author. Department of Business and Management Science, NHH Norwegian School of Economics. Helleveien 30, 5045 Bergen, Norway. \texttt{hakon.otneim@nhh.no}] and Dag Tjøstheim^[University of Bergen \newline {Acknowledgements:} The authors gratefully acknowledge the input of an associate editor as well as two anonymous referees, whose comments have greatly improved this paper.]
output: 
  bookdown::pdf_document2:
      dev: tikz
      keep_tex: yes
      number_sections: yes
      toc: no
      includes:
          in_header: latex_preamble.tex
classoption: letterpaper
geometry: margin = 1in
bibliography: "bibliography.bib"
biblio-style: "apalike"
link-citations: yes
indent: yes
fontsize: 12pt
abstract: > 
   It is well known in econometrics and other fields that the dependence structure for jointly Gaussian variables can be fully captured using correlations, and that the \textit{conditional} dependence structure in the same way can be described using partial correlations. The partial correlation does not, however, characterize conditional dependence in many non-Gaussian populations. This paper introduces the local Gaussian partial correlation (LGPC), a new measure of conditional dependence. It is a local version of the partial correlation coefficient that characterizes conditional dependence in a large class of populations. It has some useful and novel properties besides: The LGPC reduces to the ordinary partial correlation for jointly normal variables, and it distinguishes between positive and negative conditional dependence. Furthermore, the LGPC can be used to study departures from conditional independence in specific parts of the distribution. We provide several examples of this, both simulated and real, and derive estimation theory under a local likelihood estimation framework. Finally,  we indicate how the LGPC can be used to construct a powerful test for conditional independence, which, for example, can be used to detect nonlinear Granger causality in time series. \newline \textbf{Keywords:} \emph{Conditional dependence, conditional independence test, Granger causality}
---

>>>>

---
title: "The locally Gaussian partial correlation"
author: \footnote{{Acknowledgements:} The authors gratefully acknowledge the input of an associate editor as well as two anonymous referees, whose comments have greatly improved this paper.}
output: 
  bookdown::pdf_document2:
      dev: tikz
      keep_tex: yes
      number_sections: yes
      toc: no
      includes:
          in_header: latex_preamble.tex
classoption: letterpaper
geometry: margin = 1in
bibliography: "bibliography.bib"
biblio-style: "apalike"
link-citations: yes
indent: yes
fontsize: 12pt
abstract: > 
   It is well known in econometrics and other fields that the dependence structure for jointly Gaussian variables can be fully captured using correlations, and that the \textit{conditional} dependence structure in the same way can be described using partial correlations. The partial correlation does not, however, characterize conditional dependence in many non-Gaussian populations. This paper introduces the local Gaussian partial correlation (LGPC), a new measure of conditional dependence. It is a local version of the partial correlation coefficient that characterizes conditional dependence in a large class of populations. It has some useful and novel properties besides: The LGPC reduces to the ordinary partial correlation for jointly normal variables, and it distinguishes between positive and negative conditional dependence. Furthermore, the LGPC can be used to study departures from conditional independence in specific parts of the distribution. We provide several examples of this, both simulated and real, and derive estimation theory under a local likelihood estimation framework. Finally,  we indicate how the LGPC can be used to construct a powerful test for conditional independence, which, for example, can be used to detect nonlinear Granger causality in time series. \newline \textbf{Keywords:} \emph{Conditional dependence, conditional independence test, Granger causality}
---

### 2020-09-18 lgpc_JBES_blinded.tex
Created blinded versions and zipped extra material

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
  letterpaper]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The locally Gaussian partial correlation},
  pdfauthor={Håkon Otneim and Dag Tjøstheim},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin = 1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}
\usepackage{sectsty}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{bm}

@@

\title{The locally Gaussian partial correlation}
\author{Håkon Otneim\footnote{Corresponding author. Department of Business and Management Science, NHH Norwegian School of Economics. Helleveien 30, 5045 Bergen, Norway. \texttt{hakon.otneim@nhh.no}} and Dag Tjøstheim\footnote{University of Bergen \newline {Acknowledgements:} The authors gratefully acknowledge the input of an associate editor as well as two anonymous referees, whose comments have greatly improved this paper.}}
\date{}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
  letterpaper]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The locally Gaussian partial correlation},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin = 1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}
\usepackage{sectsty}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{bm}

@@

\title{The locally Gaussian partial correlation}
\author{\footnote{{Acknowledgements:} The authors gratefully acknowledge the input of an associate editor as well as two anonymous referees, whose comments have greatly improved this paper.}}
\date{}
